---
title: "Distributions"
output: html_notebook
---

```{r}
library(prob)
library(janitor)
library(tidyverse)
library(skimr)
```

1. 
```{r}
phones <- read_csv("data/20190928-items.csv") %>% 
  clean_names()
```

Explore data:
```{r}
phones %>% 
  group_by(asin) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```


```{r}
phones %>% 
  group_by(title) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

2. Find the brand with the highest numbers of phones in the dataset.
```{r}
phones %>% 
  group_by(brand) %>% 
  summarise(count = n()) %>% 
  slice_max(count)
```

3. 

```{r}
samsung_phones <- phones %>% 
  filter(brand == "Samsung")

ggplot(samsung_phones)+
  aes(x = rating)+
  #geom_density()
  geom_histogram(col = "white")
```

```{r}
skim(samsung_phones$rating)
```

```{r}

samsung_phones %>% 
  ggplot()+
  geom_line(aes(x = rating, y = dnorm(x = rating, mean = 3.57, sd = 0.685)))



three_norms <- tibble(
  x =  seq(0, 20, 0.1),
  f1_x = dnorm(x = x, mean = 10, sd = 1), 
  f2_x = dnorm(x = x, mean = 10, sd = 2),
  f3_x = dnorm(x = x, mean = 10, sd = 3)
)

```

```{r}
samsung_stats <- samsung_phones %>%
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )
samsung_stats
```


```{r}
samsung_phones %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  stat_function(
    fun = dnorm, 
    args = list(
      mean = samsung_stats$mean, 
      sd = samsung_stats$sd
    ),
    col = "red"
  )
```

- unimodal
- not skewed

It’s a reasonable fit: the distribution is roughly bell-shaped. The main problems seem to be a larger number of ratings in the tails of the distribution than are predicted by the normal, at ratings 1, 2 and 5

4.

```{r}
samsung_phones %>%
  filter(rating >= samsung_stats$mean - samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + samsung_stats$sd) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_phones))
```

```{r}
samsung_phones <- samsung_phones %>% 
  mutate(z_rating = scale(rating))
```

```{r}
samsung_phones %>% 
  filter(between(z_rating, left = -2, right = 2)) %>% 
  summarise(prop_withing_2sd = n() / nrow(samsung_phones))
```
```{r}
samsung_phones %>% 
  filter(between(z_rating, left = -3, right = 3)) %>% 
  summarise(prop_withing_3sd = n() / nrow(samsung_phones))
```
```{r}
samsung_phones %>% 
  filter(!between(z_rating, left = -2, right = 2)) %>% 
  summarise(prop_outside_3sd = n() / nrow(samsung_phones))
```
# alternative way

```{r}
samsung_phones %>%
  filter(rating >= samsung_stats$mean - 3 * samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + 3 * samsung_stats$sd) %>%
  summarise(prop_within_3sd = n() / nrow(samsung_phones))
```

5. Extention

```{r}
qqnorm(samsung_phones$rating)
qqline(samsung_phones$rating)
```
Here we see significant deviations from straight line in the tails, so the data is not very well approximated as normal.
