---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

Create three day weather forecast.

```{r}
three_days_weather <- list(
  Thursday = list(temperature = 3, chance_of_rain = 90, wind_speed = 13),
  Friday = list(temperature = 5, chance_of_rain = 86, wind_speed = 24),
  Saturday = list(temperature = 5, chance_of_rain = 88, wind_speed = 16)
)
names(three_days_weather)
```

```{r}
# The chance of rain tomorrow
three_days_weather$Thursday$chance_of_rain

#The wind speed on the day after tomorrow
three_days_weather$Friday$wind_speed

# The temperature tomorrow.
three_days_weather$Thursday$temperature

```

Find a 14 day forecast online and create a matrix of the forecasts for temperature, where each week is one row and columns correspond to days.

```{r}
fourteen_day_forcast <- matrix(c(3, 5, 5, 3, 3, 3, 4, 4, 5, 6, 3, 2, 2, 8), nrow = 2, ncol = 7, byrow = TRUE)
fourteen_day_forcast
```


```{r}
# The third day of the second week
fourteen_day_forcast[2, 3]

# The whole second week
fourteen_day_forcast[2, c(1:7)] # or [2, ]

# The first, third and fifth days in both weeks.
fourteen_day_forcast[ , c(1, 3, 4)]
```

Write a function that pre-appends “The answer is:” to any character input, while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. For anything else return “Invalid input”.

```{r}
is_char_or_num <- function(x){
  if (is.numeric(x) == TRUE){
    if (x > 10){
      print(x/10)
    } else {
      print(x)
    }
  } else {
    paste("The answer is: ", x)
  }
}

is_char_or_num(12)
is_char_or_num("Hello")
is_char_or_num(8)
```

Write a function which checks if two vectors have the same elements.

```{r}
compare_two_vectors <- function(x, y){
  if (identical(x, y) == TRUE){
    print("Exactly the same")
  } else if (all(x %in% y)) {
    print("The same")
  } else {
    print("Different")
  }
}
compare_two_vectors(c(10, 2, 3, 4, 1, 6), c(10, 2, 3, 4, 1, 6))
```
```{r}
a <- c(1, 2, 3, 4, 5)
b <- c(1, 4, 2, 3, 5)
c <- c(9, 4, 8, 1, 0)

compare_two_vectors(a, b)
compare_two_vectors(a, c)
compare_two_vectors(a, a)
```


GO THROUGH ENTENSION:

# Extension

Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

```{r}
print_multiplication_table <- function(x){
  if (x <= 12){
    outer(1:12, 1:12, "*")
  }
}
print_multiplication_table(12)
```

```{r}
tbl <- outer(1:12, 1:12, "*")
rownames(tbl) <- paste(1:12, "'s", sep="")
tbl
```
```{r}
# alternatve way 

mult_table <- matrix(nrow = 12, ncol = 12)

for (outer_factor in 1:12){
  for (inner_factor in 1:12){
    mult_table[outer_factor, inner_factor] = outer_factor * inner_factor
  }
}

mult_table
```
```{r}
# another way

mult_table <- matrix(1:12) %*% t(matrix(1:12))
mult_table
```

Write a function that takes a vector, and creates a text art histogram, like the one below.

```{r}
make_histogram <- function(x){
  
  distinct <- sort(unique(x))
  
  for (i in distinct){
    
    count <- sum(x == i)
    
    cat(i)
    cat(" | ")
    cat(rep("*", count))
    cat("\n")
  }
}
```

```{r}
numbers <- sample(1:7, size = 80, replace = TRUE) 
numbers
```

```{r}
make_histogram(numbers)
```

