---
title: "R Notebook"
output: html_notebook
---

```{r}
library(CodeClanData)
library(tidyverse)
```

```{r}
backpack <- backpack
```

2. What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}
ggplot(backpack)+
  aes(x = backpack_weight)+
  geom_histogram(colour = "white", binwidth = 5)+
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,35))+
  labs(title = "Distribution of backpack weights carried by students\n",
       x = "\nbackpack weight (lbs)",
       y = "number of students\n")

```
Description: The most common weight of student backpack is 10 pounds. The majority of students carry backpack that weight between 5 to 15 pounds.

3. Are there differences in backpack weight between sex?

```{r}
ggplot(backpack)+
  aes(x = backpack_weight, fill = sex)+
  geom_histogram(colour = "white", position = "dodge", binwidth = 5)+
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,35))
  labs(title = "Distribution of backpack weights carried by students",
       subtitle = "Differences between gender\n",
       x = "\nbackpack weight (lbs)",
       y = "number of students\n",
       fill = "gender")
```
Description: There is overall higher number of Female students carrying backpacks than Male students. 

4. What is the relationship between body weight and backpack weight?

```{r}
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight)+
  geom_point()+
  #geom_smooth(method = "lm", se = FALSE)+
  labs(title = "Relationship between body weight and backpack weight\n",
       x = "\nstudent weight (lbs)",
       y = "backpack weight (lbs)\n")
```
Description: There is no relationship between body weight and student backpack weight. Most of the students carry backpacks of weight 5 to 20 pounds with no relation to the body weight. 

5. Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

```{r}
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight, colour = sex)+
  geom_point()+
  #geom_smooth(method = "lm", se = FALSE)+
  facet_wrap(~ sex)+
  labs(title = "Relationship between body weight and backpack weight",
       subtitle = "Different between gender\n",
       x = "\nstudent weight (lbs)",
       y = "backpack weight (lbs)\n")
```
Female backpack weight is mainly centered around 5 to 15 pounds and weight is centered around 100 to 170 pounds. 
While Male backpack weight is mostly in between 5 to 20 pounds and weight around 125 to 200 pounds. 
Plot suggests that women might weigh less, but they’re not carrying less heavy backpacks

6. What proportion of students report back pain?

```{r}
backpack_bp <- backpack %>% 
  mutate(back_problems = if_else(back_problems == 1, "yes", "no"))

ggplot(backpack_bp)+
  aes(x = back_problems)+
  geom_bar(fill = "light blue")+
  labs(title = "Students reporting back problems\n",
       x = "back problems",
       y = "number of students\n")

```

Description: Approximately 1/3 of the students report back problems.

7. Is this proportion different between men and women?

```{r}
ggplot(backpack_bp)+
  aes(x = back_problems, fill = sex)+
  geom_bar()+
  facet_wrap(~sex)+
  labs(title = "Students reporting back problems\n",
       subtitle = "Difference between gender",
       x = "\nback problems",
       y = "number of students\n")
```

Description: Majority of students reporting back pain are females. 

Exceptions:

1. Does reported backpain differ across the year groups? Ignore year 0 in your data.
```{r}
backpack %>% 
  filter(year > 0) %>% 
  ggplot()+
  aes(x = year, group = back_problems, fill = back_problems) %>% 
  geom_histogram()+
  facet_wrap(~back_problems)+
  theme(legend.position = "none")
  
```
Year 3 and 4 report increase in backpack pain. There is a significant difference between year 2 and 3 in reporting back pain by students.

2. Is there a difference between how heavy the bags are in different years?

```{r}
ggplot(backpack)+
  geom_bar(aes(x = backpack_weight)) + 
  facet_wrap(~ year, ncol = 4) + 
  xlab("backpack weights")
```

Here it might be tempting to say years 3 and 4 have more people carrying heavy bags, but it actually seems more like it is the data that’s not great in the last years and we need to collect more data.