---
title: "Probability"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
```

```{r}
titanic <- read_csv("data/titanic.csv")
```

```{r}
# get initial counts of class

titanic %>%
  count(class)
```

```{r}
titanic <- titanic %>% 
  mutate(crew = if_else(class %in% c("1st", "2nd", "3rd"), FALSE, TRUE))

titanic %>%
  count(crew)
```

prob(survived | crew)
```{r}
titanic %>% 
  filter(crew == T) %>% 
  summarise(crew_count = n())
```
```{r}
titanic %>% 
  filter(crew == T & survived == "yes") %>% 
  summarise(survived_crew_count = n())
```
```{r}
211/890

titanic %>%
  filter(crew) %>%
  summarise(prob = sum(survived == "yes") / n())
```

The proportion of crew that survived was 24%.


prob(survived | passenger)
```{r}
titanic %>% 
  filter(crew == F) %>% 
  summarise(crew_count = n())
```
```{r}
titanic %>% 
  filter(crew == F & survived == "yes") %>% 
  summarise(survived_crew_count = n())
```
```{r}
500/1317

titanic %>%
  filter(!crew) %>%
  summarise(prob = sum(survived == "yes") / n())
```

The proportion of passengers that survived was 38%.

So passengers were 1.6 times more likely to have survived the disaster than crew.

1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
rolldie(1) %>% 
  filter(X1 < 5)

4/6 # 2/3
```

2. How many outcomes are there in the sample space of an experiment in which we roll three dice?

- 216 possible outcomes if we roll three dice
```{r}
6*6*6
rolldie(3) %>% 
  nrow()
```

3. 
i.
```{r}
rolldie(4) %>% 
  mutate(total = X1 + X2 + X3 + X4) %>% 
  filter(total %in% c(5, 6, 7))
```
ii.
what is the probability of this event?

```{r}
34/1296
```

# alternative way

```{r}
S <- rolldie(4)

# add total random variable
S <- S %>%
  mutate(total = rowSums(.))

# the total of the four dice equals 5, 6 or 7)
S_total_between_5_7 <- S %>%
  filter(total %in% c(5, 6, 7))

nrow(S_total_between_5_7)

# probability of this event
prob <-nrow(S_total_between_5_7) / nrow(S)
prob
```

iii.
All assumptions are fair.
We assume all outcomes in the sample space are equally likely!

4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

```{r}
permutations(
  n = 5,
  r = 5) %>% 
  as_tibble() %>% 
  nrow()

1/120
```

# alternative way

```{r}
# number of ordered arrangements:
num_outcomes <- factorial(5)

# only one of these is correctly ordered:
prob_perfect_order <- 1/num_outcomes
prob_perfect_order
```

5. I roll six dice:
i. what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right? - 1/46656

```{r}
rolldie(6)
1/46656
```
# alternative ways

```{r}
S <- rolldie(6)

prob <- 1 / nrow(S)
prob

# or
S_in_order <- S %>%
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = TRUE))
prob <- nrow(S_in_order) / nrow(S)
prob

# or
prob <- 1 / 6^6
prob
```


ii. what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
```{r}
S <- rolldie(6)

S_contains_654321 <- rolldie(6) %>% 
  filter(isin(., c(1,2,3,4,5,6), ordered = FALSE)) %>% 
  nrow()

prob <- nrow(S_contains_654321) / nrow(S)
prob
```
```{r}
factorial(6)
```

6. If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

```{r}
S <- cards(jokers = FALSE)

S <- S %>%
  mutate(black_card = suit == "Club" | suit == "Spade")

S_spades_and_black <- S %>%
  filter(black_card == TRUE & suit == "Spade")

prob_spades_and_black <- nrow(S_spades_and_black) / nrow(S)

S_black <- S %>%
  filter(black_card == TRUE)

prob_black <- nrow(S_black) / nrow(S)

prob_spades_given_black <- prob_spades_and_black / prob_black
prob_spades_given_black
```


7. Eight patients: Alice, Bob, Carmen, Dinesh, Ella, Fergus, Gianna and Henry are waiting to see a doctor, but only five appointment slots are left today.

i. In how many different orders can the doctor see patients today?

```{r}
patients <- c("Alice", "Bob", "Carmen", "Dinesh", "Ella", "Fergus", "Gianna", "Henry")

permutations(
  n = 8,
  r = 5,
  v = patients
) %>% 
  as_tibble() %>% 
  nrow()
```
```{r}
# or

8*7*6*5*4
```

ii. What is the probability of Gianna and Dinesh both being seen today?

```{r}
permutations(
  n = 8,
  r = 5,
  v = patients
) %>% 
  as_tibble() %>% 
  filter(isin(., c("Gianna", "Dinesh"))) %>% 
  nrow()
```

```{r}
2400/6720
```

îii. What is the probability of Gianna being seen before Dinesh today?

```{r}
permutations(
  n = 8,
  r = 5,
  v = patients
) %>% 
  as_tibble() %>% 
  filter(isin(., c("Gianna", "Dinesh"), ordered = TRUE)) %>% 
  nrow()
```

```{r}
1200/6720
```

iv. Carmen and Henry have now been triaged as high priority and will be seen in the first two available slots, although their order hasn’t yet been determined. In how many different orders can the patients now be seen?

```{r}
240
```

8.  If order is not significant, how many ways are there to select a 5 student quiz team from a group of 8 students?

```{r}
combinations(
  n = 8,
  r = 5
) %>% 
  as_tibble() %>% 
  nrow()
```

9. 
```{r}
combinations(
  n = 7,
  r = 4
) %>% 
  as_tibble() %>% 
  nrow()
```

```{r}
35*8
```

# alternative way

```{r}
# step 1: choose a captain. We have 8 choices
num_captains <- 8

# step 2: for each choice of captain, choose 4 remaining team members from 7 
teams <- combinations(n = 7, r = 4, v = 1:7, repeats.allowed = FALSE)
num_team_combinations <- nrow(teams)

# combinations
num_captains * num_team_combinations
```

