---
title: "Cancer Rates in NHS Borders"
output: html_notebook
---


```{r}
# packages required
library(tidyverse)
library(janitor)
library(readxl)

# data on geography codes and labels
geography_codes <- read_csv("data/geography_codes.csv") %>% 
  clean_names()

# cancer incidents by Health Board
incidence_by_health_board <- read_csv("data/incidence_by_health_board.csv") %>% 
  clean_names()
```

```{r}
# join two tables and remove variables that aren't required
cancer_rates <- left_join(incidence_by_health_board, geography_codes, by = "hb") %>% 
  filter(hb_name == "NHS Borders") %>% 
  select(-c(hb_date_enacted, hb_date_archived, country))
```

Graph to visualise all registered cancer incidence in NHS Boarders, 1994 to 2018
```{r}
cancer_rates %>% 
  group_by(year) %>% 
  filter(cancer_site == "All cancer types") %>% 
  filter(sex == "All") %>% 
  ggplot()+
  aes(x = year, y = incidences_all_ages)+
  geom_line(size = 1.5, color = "#FF9999")+
  labs(title = "All registered cancer incidences in NHS Borders, 1994 to 2018\n",
       x = "\nYear",
       y = "Incidence\n")+
  theme_linedraw()+
  scale_x_continuous(breaks = c(1994, 1998, 2002, 2006, 2010, 2014, 2018))
```

Total number of most common new cancer registrations in NHS Borders, 2008 to 2018
```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>%
  group_by(year, cancer_site) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  filter(sex == "All") %>% 
  filter(incidences_all_ages > 50) %>% 
  ggplot()+
  aes(x = year, y = incidences_all_ages, fill = cancer_site)+
  geom_point(size = 3, shape = 21)+
  scale_fill_brewer(palette = "Set1")+
  labs(fill = "Diagnosed cancer",
       title = "Most common new registered incidences of cancer by year, 2008 to 2018",
       subtitle = "Exceeding 50 incidence per year\n",
       x = "Year",
       y = "Incidence")+
  theme_linedraw()+
  scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))

```

Most common incidences of cancer over ten year period, 2008 to 2018
```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  group_by(cancer_site) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  filter(sex == "All") %>% 
  filter(incidences_all_ages > 50) %>% 
  ggplot()+
  aes(reorder(x = cancer_site, incidences_all_ages), y = incidences_all_ages)+
  geom_col(fill = "#FF9999")+
  coord_flip()+
  labs(title = "Most common incidences of cancer by cancer site",
       subtitle = "Exceeding 50 incidences per year\n",
       x = "Cancer site",
       y = "Incidence")+
  theme_linedraw()
```


Sum of all new cancer registrations by cancer site between 2008 to 2018 in NHS Borders
```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  group_by(cancer_site) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  filter(sex == "All") %>% 
  filter(incidences_all_ages <= 50) %>% 
  #distinct(cancer_site) %>% 
  ggplot()+
  aes(reorder(x = cancer_site, incidences_all_ages), y = incidences_all_ages)+
  geom_col(fill = "#FF9999")+
  coord_flip()+
  labs(y = "Incidences",
       x = "Cancer site",
       title = "Cancer incidences by cancer site",
       subtitle = "with 50 and less registered incidences per year")+
  theme_bw()
ggsave("p.png")
```

Annual difference in instances of all ages and sex.
```{r}
cancer_rates %>% 
  group_by(cancer_site) %>% 
  filter(cancer_site == "All cancer types") %>% 
  filter(sex == "All") %>% 
  mutate(volume = 
           incidences_all_ages - lag(incidences_all_ages, default = incidences_all_ages[1])) %>% 
  arrange(desc(volume)) %>% 
  select(volume, year)
```

Number of instances of particular cancer type over ten years
```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  filter(cancer_site == "Malignant melanoma of the skin") %>% 
  filter(sex == "All") %>% 
  summarise(sum(incidences_all_ages))
```
Amount of registered cancer sites in ten year period
```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  distinct(cancer_site)
  
  
```

```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  distinct(cancer_site)
```

```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  filter(incidences_all_ages <= 50) %>% 
  distinct(cancer_site)
```

Check up code on above calculations
```{r}
cancer_rates %>% 
  filter(year %in% 2008:2018) %>% 
  filter(!cancer_site == "All cancer types") %>% 
  filter(sex == "All") %>% 
  group_by(cancer_site) %>% 
  summarise(total = sum(incidences_all_ages)) %>% 
  arrange(desc(total))
```

