---
title: "SQL & Programming"
output: html_notebook
---
3.
```{r}
source("~/acme_credentials.R")
```

```{r}
library(RPostgres)
library(tidyverse)
```

```{r}
db_connection = dbConnect(
  drv = Postgres(), 
  user = acme_username,
  password = acme_password,
  dbname = acme_database,
  host = acme_host,
  port = acme_port,
  sslmode = 'require',
  bigint = "numeric")
```

```{r}
rm(acme_database, acme_host, acme_password, acme_port, acme_username)
```

```{r}
dbListTables(conn = db_connection)
```

```{r}
dbListFields(conn = db_connection, name = "employees")
```

4. The records of all employees, adding a field full_name containing the first_name and last_name of each employee separated by a space.

```{sql, connection = db_connection}
SELECT 
  id,
  CONCAT(first_name,' ' ,last_name) AS full_name,
  email,
  team_id
FROM employees
```
5. All employees records again, but this time listing their team name (as team_name), rather than their team_id. 

```{sql, connection = db_connection}
SELECT 
  e.id,
  e.first_name,
  e.last_name,
  e.email,
  t.name AS team_name
FROM employees AS e LEFT JOIN teams AS t
ON e.team_id = t.id
```
6. A breakdown of the number of employees per team. The table is ordered by decreasing number of employees. Include all teams, even if they have no employees.

```{sql, connection = db_connection}
SELECT
  team_id,
  COUNT(id) AS num_employees
FROM employees
GROUP BY team_id
ORDER BY COUNT(id) DESC
```
Alternative way

```{sql, connection = db_connection}
SELECT
  t.name AS team_name,
  COUNT(e.id) AS num_employees
FROM employees AS e RIGHT JOIN teams AS t
ON e.team_id = t.id
GROUP BY t.id
ORDER BY COUNT(e.id) DESC
```

7. The employees that the managment cannot send an email because their email addresses may be invalid (in that they don’t contain an ‘@’ symbol), or they may not have a stored email address at all.

```{sql, connection = db_connection}
SELECT *
FROM employees
WHERE email NOT LIKE '%@%' OR email IS NULL
```

Alternative way

```{sql, connection = db_connection}
SELECT * 
FROM employees
WHERE email !~ '[a-zA-Z0-9_.-]+@[a-zA-Z0-9_.-]+' OR email IS NULL
```

9. The employees that provided the most recent feedback with the date of this feedback.

```{sql, connection = db_connection}
SELECT
  e.id,
  e.first_name,
  e.last_name,
  f.date
FROM employees AS e INNER JOIN feedbacks AS f
ON e.id = f.employee_id
ORDER BY date DESC
```
10. Team with the highest number of employees with only a single initial for their first_names.

```{sql, connection = db_connection}
SELECT
  t.name,
  COUNT(e.id) AS num_employees
FROM employees AS e INNER JOIN teams AS t
ON e.team_id = t.id
WHERE e.first_name LIKE '_'
GROUP BY t.name
ORDER BY COUNT(e.id) DESC
LIMIT 1
```
Alternative way

```{sql, connection = db_connection}
WITH counts_only_first_initial(team_name, num_only_first_initial) AS (
  SELECT 
    t.name,
    COUNT(e.id)
  FROM employees AS e LEFT JOIN teams AS t
  ON e.team_id = t.id
  WHERE e.first_name LIKE '_'
  GROUP BY t.name
)
SELECT 
  team_name,
  num_only_first_initial
FROM counts_only_first_initial
WHERE num_only_first_initial = (
  SELECT 
    MAX(num_only_first_initial)
  FROM counts_only_first_initial
)
```

11. Combining SQL and R coding - basic sentiment analysis

```{r}
positive_words <- c("great", "good", "happy", "content", "fulfilled", "fulfilling", "positive", "positives", "cheerful", "nice", "ambition", "ambitions", "ambitious", "progress", "joy", "fast", "achieve", "achieving", "achievements", "best", "better",  "unlimited", "pros", "competitive", "opportunity", "advancement", "friendly", "smart", "interesting", "challenging", "grow", "caring", "balance", "helpful", "advantage", "learning", "ethical", "responsible", "responsibilities")

negative_words <- c("worse", "worsening", "worst", "slow", "sluggish", "impatient", "unfulfilled", "unfulfilling", "negative", "bad", "limited", "limit", "limitations", "cons", "frustrated", "frustrating", "frustration", "dull", "boring", "politics", "political", "ignore", "ignored", "ignoring", "sad", "depressed", "depressing", "nasty", "stress", "stresses", "stressed", "stressful", "horrid", "horrible", "terrible", "deadline", "deadlines", "uncaring", "remote", "difficult", "difficulties", "bully", "bullying")
```

```{sql, connection = db_connection}
SELECT
  e.*, 
  f.message,
  f.id AS message_id,
  t.name AS team_name
FROM employees AS e LEFT JOIN teams AS t
ON e.team_id = t.id INNER JOIN feedbacks AS f
ON f.employee_id = e.id
```
```{r}
employees_teams_feedbacks <- dbGetQuery(conn = db_connection, statement = "SELECT
  e.*, 
  f.message,
  f.id AS message_id,
  t.name AS team_name
FROM employees AS e LEFT JOIN teams AS t
ON e.team_id = t.id INNER JOIN feedbacks AS f
ON f.employee_id = e.id")
```

```{r}
# clean the messages
employees_teams_feedbacks <- employees_teams_feedbacks %>%
  mutate(
    message = str_to_lower(message),
    message = str_remove_all(message, "[[:punct:]]"), # remove punctuation
    message = str_squish(message), # all internal whitespace goes to single space
    message = str_trim(message) # leading and trailing whitespace removed
  )
```

```{r}
# create a tibble containing positive and negative words and their scores
sentiment_words <- tibble(word = positive_words,
                          score = 1) %>% 
  bind_rows(tibble(word = negative_words, score = -1))

sentiment_words
```
```{r}
# split messages to words, unnest words to individual rows, 
# join to sentiment_words tibble matching on word, group word scores for each message
# and then sum to get total message scores
message_scores <- employees_teams_feedbacks %>%
  mutate(word = str_split(message, " ")) %>% # split messages to words
  unnest_longer(word) %>% # unnest words to individual rows
  inner_join(sentiment_words, by = "word") %>%  #join to sentiment_words tibble matching on word
  group_by(message_id) %>%
  summarise(score = sum(score))
```

```{r}
message_scores
```
```{r}
# join message scores back to original tibble matching on message_id
# fill in any missing scores as zero.
employees_teams_feedbacks_scores <- employees_teams_feedbacks %>%
  left_join(message_scores, by = "message_id") %>% 
  mutate(score = replace_na(score, 0)) 

employees_teams_feedbacks_scores
```
Team with the employees making the most positive comments.

```{r}
employees_teams_feedbacks_scores %>%
  filter(score == max(score)) %>%
  select(team_name)
```
Team with the employees making the most negative comments.

```{r}
employees_teams_feedbacks_scores %>%
  filter(score == min(score)) %>%
  select(team_name)
```


```{r}
dbDisconnect(conn = db_connection)
```

