---
title: "R Notebook"
output: html_notebook
---

```{r}
library(CodeClanData)
library(tidyverse)
```

```{r}
backpack <- backpack
```

```{r}
a <- unique(backpack$backpack_weight)
max(a)
min(a)
```

2.

```{r}
ggplot(backpack)+
  aes(x = backpack_weight)+
  geom_histogram(colour = "white", binwidth = 5)+
  labs(title = "Distribution of backpack weights carried by students\n",
       x = "\nbackpack weight (lbs)",
       y = "number of students\n")

```
Description: The most common weight of student backpack is 10 pounds. The majority of students carry backpack that weight between 5 to 15 pounds.

3. 
```{r}
ggplot(backpack)+
  aes(x = backpack_weight, fill = sex)+
  geom_histogram(colour = "white", position = "dodge", binwidth = 5)+
  labs(title = "Distribution of backpack weights carried by students",
       subtitle = "Differences between gender\n",
       x = "\nbackpack weight (lbs)",
       y = "number of students\n",
       fill = "gender")
```
Description: There is overall higher number of Female students carrying backpacks than Male students. 

4. 

```{r}
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight)+
  geom_point()+
  geom_smooth()+
  labs(title = "Relationship between body weight and backpack weight\n",
       x = "\nbody weight (lbs)",
       y = "backpack weight (lbs)\n")
```
Description: There is no relationship between body weight and student backpack weight. Most of the students carry backpacks of weight 5 to 20 pounds with no relation to the body weight. 

5. 

```{r}
ggplot(backpack)+
  aes(x = body_weight, y = backpack_weight, colour = sex)+
  geom_point()+
  geom_smooth()+
  facet_wrap(~ sex)+
  labs(title = "Relationship between body weight and backpack weight",
       subtitle = "Different between gender\n",
       x = "\nstudent weight (lbs)",
       y = "backpack weight (lbs)\n")
```
Female backpack weight is mainly centered around 5 to 15 pounds and weight is centered around 100 to 170 pounds. 
While Male backpack weight is mostly in between 5 to 20 pounds and weight around 125 to 200 pounds. 

6. 

```{r}
ggplot(backpack)+
  aes(x = back_problems)+
  geom_bar(fill = "")+
  coord_polar()
```

7.

```{r}
ggplot(backpack)+
  aes(x = back_problems, fill = sex)+
  geom_bar()
```

8.

```{r}

```

