---
title: "Model Building"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(ggfortify)
library(modelr)
library(relaimpo)
```

```{r}
avocado_raw <- read_csv("data/avocado.csv")
```

# Data Cleaning

- check missing values
- change variable class if required
- check redundant predictors

```{r}
avocado <- avocado_raw %>% 
  clean_names() %>% 
  mutate(month = as_factor(format(date, "%m"))) %>% 
  select(-date) %>% 
  mutate(year = as_factor(year)) %>% 
  mutate(organic = ifelse(type == "organic", TRUE, FALSE)) %>% 
  select(-c(type, region, x1, small_bags, large_bags, x_large_bags)) 
```


```{r}
#summary(avocado)

alias(lm(average_price ~ ., data = avocado))
```

```{r}
avocado_numeric <- avocado %>%
  select_if(is.numeric)

avocado_nonnumeric <- avocado %>%
  select_if(function(x) !is.numeric(x))

avocado_nonnumeric$average_price <- avocado$average_price

ggpairs(avocado_numeric)

```

```{r}
ggpairs(avocado_nonnumeric)
```
Looking at the graphs I chose to compare models for features organic, month, x 4046, total volume, total bags
```{r}
model_organic <- lm(average_price ~ organic, data = avocado)
summary(model_organic)
autoplot(model_organic)
```

# Second predictor

```{r}
avocado_residuals <- avocado %>%
  add_residuals(model_organic) %>%
  select(-c("average_price", "organic"))

ggpairs(avocado_residuals)

```

Comparining month and year
```{r}
model_month <- lm(average_price ~ organic + month, data = avocado)
summary(model_month)
autoplot(model_month)

anova(model_organic, model_month)
```

# Third predictor

```{r}
avocado_residuals <- avocado %>%
  add_residuals(model_month) %>%
  select(-c("average_price", "organic", "month"))

ggpairs(avocado_residuals)
```

Looking at total_volume and year
```{r}
model_year <- lm(average_price ~ organic + month + year, data = avocado)
summary(model_year)
```
# Fourth predictor

```{r}
avocado_residuals <- avocado %>%
  add_residuals(model_year) %>%
  select(-c("average_price", "organic", "month", "year"))

ggpairs(avocado_residuals)
```

```{r}
model_total <- lm(average_price ~ organic + month + year + total_volume, data = avocado)
summary(model_total)
```

```{r}
anova(model_year, model_total)
```

The final model includes average_price, organic, month, year, total_volume.

On average organic avocado is 0.49 cents more expensive compared with conventional, with all other predictors held constant.

One cent increase in price decrease the total volume of sold avocados by 5.576e-09, with all other predictors held constant.

The volume doesnt have signifficant inpact on the average price.

The avocados are on average 0.08 cents more expensive in June compared with other months, with all other predictors held constant.

The avocados are on average 0.06 cents cheaper in February compared with other months, with all other predictors held constant.

The average price of avocados was 0.10 cent higher in year 2017 compared to other years, with all other predictors held constant.

```{r}
calc.relimp(model_total, type = "lmg", rela = TRUE)
```

Proportion of variance explained by model: 47.62%.
So, we see by this measure that organic is most important (accounting for 76% of r2), followed by month (12%), then year (7%), and finally the total volume (4%).

