---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(tidyverse)
library(yardstick)

titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

# Data cleaning

```{r}
titanic_set <- titanic_set %>% 
  select(-c("X1", "passenger_id", "name", "ticket", "cabin", "fare")) %>% 
  filter(!is.na(survived)) %>% 
  filter(!is.na(age)) %>% 
  mutate(sex = as_factor(sex),
         age = as.factor(if_else(age <= 16, "Child", "Adult")),
         pclass = factor(pclass, levels = c(3,2,1), labels = c("Lower", "Middle", "Upper")),
         survived = factor(survived, levels = c(0,1), labels = c("Died", "Survived")),
         embarked = as.factor(embarked)) %>% 
  na.omit()

summary(titanic_set)
```
# Decision tree

```{r}
# get how many rows are in our dataset
n_data <- nrow(titanic_set)

# how many rows test is going to take
test_index <- sample(1:n_data, size = n_data * 0.2)

# make our datasets
titanic_test <- slice(titanic_set, test_index)
titanic_train <- slice(titanic_set, -test_index)
```

```{r}
titanic_fit <- rpart(survived ~ ., # chooseing what to compare to
                     data = titanic_train,
                     method = "class")

titanic_fit
```

```{r}
rpart.plot(titanic_fit, digits = 2, yesno = 2)
```
If character's sex is male then there is 42% probability of death. 
If character's ticket class is Middle or Lower then there is 21% probability of death.
If character's is adult then there is 16% probability of death.
If character is a child and has three or more siblings on board then they probability of death is 37%.
If children with siblings less than three and with at least one parent on board has probability of survival is 61%
# Calculate accuracey

```{r}
titanic_test_pred <- titanic_test %>% 
  add_predictions(titanic_fit, type = "class")

conf_mat <- titanic_test_pred %>% 
  conf_mat(truth = survived, estimate = pred)

conf_mat
```

# Three measures to see how well the model is predicting

```{r}
titanic_test_pred %>% 
  accuracy(truth = survived,
           estimate = pred)

```
82% of predictions are correct.

```{r}
titanic_test_pred %>% 
  sensitivity(truth = survived,
              estimate = pred)
```
```{r}
titanic_test_pred %>% 
  specificity(truth = survived,
              estimate = pred)
```

