---
title: "Manual Model Building"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
```

```{r}
kc_housing <- read_csv("data/kc_house_data.csv")
```

# Tidy Data

```{r}
# removing some variables
kc_housing <- kc_housing %>% 
  select(-c("date", "id", "sqft_living15", "sqft_lot15", "zipcode"))
```

```{r}
# converted yr_renovated into a renovated logical variable, indicating whether the property had ever been renovated or no
# converted grade into a above average construction conditions and below
kc_housing <- kc_housing %>% 
  mutate(renovated = ifelse(yr_renovated == 0, 0, 1),
         above_avg_constraction = ifelse(grade %in% c(1:6), 0, 1)) %>% 
  select(-c(yr_renovated, grade))
```

```{r}
# remove variables that lead to an alias
kc_housing <- kc_housing %>% 
  select(-c(sqft_basement, sqft_above))

alias(lm(price ~ ., data = kc_housing))

# check for any missing values
summary(kc_housing)
```

# Systematically build a regression model 

```{r}
kc_housing_numeric <- kc_housing %>%
  select(price, bedrooms, bathrooms, sqft_living, sqft_lot, floors, yr_built, lat, long)

kc_housing_nonnumeric <- kc_housing %>%
  select(price, waterfront, view, condition, renovated, above_avg_constraction)

ggpairs(kc_housing_numeric)
```

```{r}
ggpairs(kc_housing_nonnumeric)
```
price and sqft_living
```{r}
kc_housing %>%
  ggplot(aes(x = sqft_living, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
There seems to be a linear relationship: as the sqft_living goes up, the price goes up

```{r}
# Goodness-of-fit metrics

model_p_living <- lm(price ~ sqft_living, data = kc_housing)

autoplot(model_p_living)
```
1. residuals are randomly scattered around zero
2. most of the points are on the straight line
3. the line doesn't stay constant on a positive value - fail

price and bathrooms
```{r}
kc_housing %>%
  ggplot(aes(x = bathrooms, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
There seems to be a linear relationship: more bathrooms, the price goes up

```{r}
# Goodness-of-fit metrics

model_p_bathrooms <- lm(price ~ bathrooms, data = kc_housing)

autoplot(model_p_bathrooms)
```
Price and latitude
```{r}
kc_housing %>%
  ggplot(aes(x = lat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
model_p_lat<- lm(price ~ lat, data = kc_housing)

autoplot(model_p_lat)

summary(model_p_lat)
```

