---
title: "Hypothesis tests & probability - homework"
output: html_notebook
---
----------------------------------
1.1 Hypothesis testing - practical
----------------------------------
```{r}
library(tidyverse)
library(infer)
data(msleep)
```

#1. 
Jabberwockies sleep for around 7 hours a night, on average. Perform an appropriate statistical test to determine whether the mean sleep_total in the sampled population of animal types differs from the typical value for jabberwockies.

H0: μ(sleep_total) = 7
Ha: μ(sleep_total) ≠ 7

significance level = 0.05 (5%)

```{r}
# generate the null distribution, assuming Null Hypothesis is true

null_distribution <- msleep %>% 
  specify(response = sleep_total) %>% 
  hypothesise(null = "point", mu = 7) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")

# value of observed statistic, the mean(sleep_total)

observed_stat <- msleep %>% 
  summarise(mean_sleep_total = mean(sleep_total))

# Visualise the null distribution

null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat$mean_sleep_total, direction = "both")

# Calculate p-value

p_value <- null_distribution %>% 
  get_p_value(obs_stat = observed_stat$mean_sleep_total, direction = "both")
```

# Conclusion:

p_value < 0.001 and significance level = 0.05.
0.001 < 0.05 (p_value < significance level)

Reject the Null.

There is enough evidence to believe that the avarage sleep hours (mean sleep_total) 
in the sampled population of animals differs from the typical (7 hours) value for jabberwockies. 

# 2. 
Perform an appropriate statistical test to determine whether omnivores sleep for significantly longer than herbivores, on average.

- two-sample independent hypothesis test

α = 0.05

H0: μ sleep_total(omnivores) − μ sleep_total(herbivores) = 0
Ha: μ sleep_total(omnivores) − μ sleep_total(herbivores) > 0

```{r}
# selecting the data we want to use

msleep_filtered <- msleep %>% 
  select(vore, sleep_total) %>% 
  filter(vore %in% c("omni", "herbi"))

# visualise the distributions and check the sizes of both samples

msleep_filtered %>%
  group_by(vore) %>%
  summarise(n = n())

msleep_filtered %>%
  ggplot(aes(y = sleep_total, x = vore)) +
  geom_boxplot()

# using permutation to generate the null distribution

null_distribution <- msleep_filtered %>% 
  specify(sleep_total ~ vore) %>% # testing whether sleep_total varies with vore
  hypothesize(null = "independence") %>% # the null hypothesis is there is no relationship
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

# calculate observed statistic 

observed_stat <- msleep_filtered %>% 
  specify(sleep_total ~ vore) %>%
  calculate(stat = "diff in means", order = c("omni", "herbi"))

# visualise the null distribution and where the observed statistic lies on the distribution

null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "greater")

# calculate p-value

p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "greater")

p_value
```
# Conclusion

p-value (0.111) > α (0.05) 
 We fail to reject the Null Hypothesis.

We can conclude that there is not enough evidence in our data to suggest that omnivores sleep on average for significantly longer than herbivores.


# 3
Perform an appropriate statistical test to determine whether the proportion of domesticated animal types in the population of animal types is greater than 5%.

H0: π(is_domesticated == "yes") = 0.05
Ha: π(is_domesticated == "yes") > 0.05

α = 0.05

```{r}
# creating an is_domesticated variable for the analysis

msleep_tidy <- msleep %>% 
  filter(!is.na(conservation)) %>% 
  mutate(is_domesticated = if_else(conservation == "domesticated", "yes", "no"))


# generating null sampling distribution, using simulation

null_distribution <- msleep_tidy %>% 
  specify(response = is_domesticated, success = "yes") %>% 
  hypothesise(null = "point", p = 0.5) %>% 
  generate(reps = 5000, type = "simulate") %>% 
  calculate(stat = "prop")

# calculate point estimate of our sample population
observed_stat <- msleep_tidy %>% 
  specify(response = is_domesticated, success = "yes") %>% 
  calculate(stat = "prop")


# visualise sampling distribution

null_distribution %>% 
  visualise(bins = 30)+
  shade_p_value(obs_stat = observed_stat, direction = "greater")

# calculate p-value

p_value <- null_distribution %>% 
  get_p_value(obs_stat = observed_stat, direction = "greater")
```

# Conclusion

p_value (1) > α (0.05)

Fail to reject the null hypothesis.

There is not sufficient evidence to believe that the proportion of domesticated animal types in the population of animal types is greater than 5%.
