---
title: "Simple Linear Regression"
output: html_notebook
---

```{r}
library(tidyverse)

project_management <- read_csv("data/project_management.csv")
```

estimated_length - contains the estimated length of a building job in days
actual_length - contains the actual recorded length of the job in days

We are interested in determining the accuracy of the job estimations made by the company using simple linear regression,

actual_length as the dependent variable - y
estimated_length as the independent variable - x

1. Plot the data, taking estimated_length as the independent variable and actual_length as the dependent variable.

```{r}
project_management %>% 
  ggplot(aes(x = estimated_length, y = actual_length))+
  geom_point()
```

2. Calculate the correlation coefficient of estimated_length and actual_length and interpret the value you obtain.
```{r}
project_management %>% 
  summarise(cor(estimated_length, actual_length))
```

Correlation coefficient is very strong.

3. Perform a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. Save the model object to a variable.
```{r}
model <- lm(formula = actual_length ~ estimated_length, data = project_management)

model
```

4. Interpret the regression coefficient of estimated_length (i.e. slope, gradient) you obtain from the model. How do you interpret the r2 value reported by the model?

```{r}

```

